---
title: "GEOG-533 Lab 9"
author: Abraham Redie
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package **UScensus2010county** to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the **plot** function.
```{r}
if(!require(UScensus2010)) install.packages("UScensus2010")
library(UScensus2010)
if(!require(UScensus2010county)) install.county("osx")
if(!require(UScensus2010tract)) install.tract("osx")
library(UScensus2010county)
library(UScensus2010tract)
data(new_york.county10)
shp <- new_york.county10
plot(shp)

```

### Question 1(b)	
Plot a map of New York counties using the **qtm** function.
```{r}
library(tmap, tmap_options(max.categories = 62))
qtm(shp, "NAME10" )
```


### Question 1(c)	
How many counties in New York State?
```{r}
nrow(shp)
```

### Question 1(d)	
What’s the 3-digit **fips** code of Broome County?
```{r}
Broomes <- shp[shp$NAME10 == "Broome", c("fips","county")]
Broomes$county
```

### Question 1(e)	
Compute descriptive statistics of the population column (**P0010001**), including total, minimum, maximum, mean, median, and skewness. 
```{r}
cat("total",sum(shp$P0010001), "\n")
cat("Minimum",min(shp$P0010001), "\n")
cat("Maximum",max(shp$P0010001), "\n")
cat("Mean",mean(shp$P0010001), "\n")
cat("Median",median(shp$P0010001), "\n")

```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(shp$P0010001)
boxplot(shp$P0010001)
```


## Question 2
Use R package **UScensus2010tract** to complete the following tasks:    (20 pt.)
```{r}
library(UScensus2010tract)
```

### Question 2(a)	
Plot a map of New York census tracts using the **plot** function.
```{r}
data(new_york.tract10)
Tract <- new_york.tract10
plot(Tract)
```

### Question 2(b)
Compute the total population based on census tracts.
```{r}
cat("Total population:",sum(Tract$P0010001), "\n")
cat("Minimum population:",min(Tract$P0010001), "\n")
cat("Maximum population:",max(Tract$P0010001), "\n")
cat("Mean population:",mean(Tract$P0010001), "\n")
cat("Median population:",median(Tract$P0010001), "\n")
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
Brom_Tract <- Tract[Tract$county == "007", ]
plot(Brom_Tract)
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
cat("Total Population is:",sum(Brom_Tract$P0010001))  
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(Brom_Tract$P0010001)
boxplot(Brom_Tract$P0010001)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r}
library(rgdal)
parts <- Brom_Tract[,1:5] 
T_Popn <- sum(parts$P0010001) 
parts$Population_ratio <- parts$P0010001/T_Popn
writeOGR(parts,dsn = ".",layer = "Broome",driver = "ESRI Shapefile", overwrite_layer = TRUE)
shpf <- readOGR(dsn = ".",layer = "Broome" )
plot(shpf)
```


## Question 3

Use R packages **raster** and **ncdf4** to complete the following tasks:     (20 pt.)
```{r}
library(raster)
library(ncdf4)
```

### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
Raster <- "E:\\R_class\\Master_lab\\Lab_09\\NDVI.nc"
Rast <- brick(Raster)
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
cat("Number of rows:",nrow(Rast), "\n")
cat("Number of column:",ncol(Rast), "\n")
cat("Number of cells:",ncell(Rast),"\n")
cat("Nmber of bands:",nbands(Rast), "\n")
cat("Resolution:", res(Rast), "\n")
extent(Rast)
cat("Bounding Box:",bbox(Rast), "\n")
cat("Projection:",projection(Rast), "\n")

```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
Mn_NDVI <- mean(Rast)
plot(Mn_NDVI, main = " Mean NDVI")
writeRaster(Mn_NDVI, filename = "Mean NDVI.tif", overwrite = TRUE)
Mx_NDVI <- max(Rast)
plot(Mx_NDVI, main = "Maximum NDVI")
writeRaster(Mx_NDVI, filename = "Max NDVI.tif", overwrite = TRUE)
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
plot(Rast,13:24)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r}
hist(Rast,13:24)
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r}
#the click function did not work.
plot(Rast,7) 
values <- click(Rast, n=7, xy=TRUE) 
values <- Rast[40,40]
plot(as.vector(values), type="b")

```



